---
title: "dnds analysis symbionts vs mitochondria"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd('./')
setwd("/Users/maeperez/Desktop/VesicSymb_Evolution/dowstream_analyses_and_data")
library('ggplot2')
library(RColorBrewer)
library(dplyr)
library(reshape2)
library("data.table")
library(gridExtra)
library(grid)
R.version
file.path(R.home("bin"), "R")
Sys.getenv('R_HOME')
```

# Introduction
dN and dS were calculated using the ML method in python
Goldman and Yang (1994): http://mbe.oxfordjournals.org/content/11/5/725

Gigas = Clade I, Ruthia = Clade II
#Import data

```{r import data }
path='./'
wide<-read.csv('./dsNuclMito_results.txt',header=T,sep='\t')
Pair=as.factor(wide$X)
wide$X<-NULL
wide$Pair=Pair
# wide=as.data.table(wide)
wide %>% head()
wide$Groups=factor(wide$Groups, level=c('F-F','F-V','F-R','V-V','R-V','R-R'))
known_isolates=c('diagonalis',  'rectimargo', 'gigas1', 'phaseoliformis', 'extenta', 'soyoae1', 'soyoae2', 'gigas2', 'pliocardia', 'fausta',  'southwardae', 'pacifica', 'marissinica')

wide$sameIsolate=as.numeric(strsplit(as.character(wide$Pair),split = "-") %>% sapply( ., "[", 2 ) %in% known_isolates & strsplit(as.character(wide$Pair),split = "-") %>% sapply( ., "[", 1 ) %in% known_isolates)
str(wide)

long=data.table::melt(setDT(wide), id.vars = c('Groups','Pair','sameIsolate'),
          measure.vars = list(c(1,4), c(2,5), c(3,6), c(12,13),c(8,10),c(9,11)),
          variable.name='genome',
          value.name=c('dnds','ds','dn','f84','ts','tv'))

# long=data.table::melt(setDT(wide),id.vars=c('Groups','Pair','sameIsolate'),
#             measure.vars = patterns("dnds_", "ds_","dn_","transition_","transversion_","f84_"),
#             variable.name = 'genome',
#             value.name=c('dnds','ds','dn','ts','tv','f84'),
#             variable.factor = T)

str(long)
long$genome= ifelse(long$genome == 1,'mito','symb')
summary(long)
long$Pair=as.factor(long$Pair)
long$genome=as.factor(long$genome)
long %>% tail()
```

# Check for substitution saturation

There doesnt seem to be saturation. However FL-symb pairs are close to saturation
```{r}
# TSTV vs dS in mito and symb

ggplot(long, aes(color=Groups))+
  geom_point(aes(x=f84,y=ts,shape='ts'))+
  geom_point(aes(x=f84,y=tv,shape='tv'))+
  facet_wrap(genome~., scales='free')+
  ylab('Ts and Tv')
# TS/TV free-living close to equal; indicates saturation


ggplot(long)+
  geom_point(aes(x=f84,y=ts,color='ts'))+
  geom_point(aes(x=f84,y=tv,color='tv'))+
  facet_wrap(genome~., scales='free')

```

# Ds in mito vs symbionts
The symbionts evolve more slowly than the mitochondria --> Recombination offers a degree of sex mitochondria dont have.
Amongst the symbionts, the Ruthia does not evolve faster than Gigas but evolution acceleration happened on the branch leading to gigas.

```{r}
wide %>% head

colors=list(Ve='#606E96',Ru='#68C47F',FL='#E1A79E')

### Symbiont vs mitochondrial divergence highlighting pairs of genomes issued from the same individual hosts, respectively ###

ds_plot=ggplot(subset(wide, Groups %in% c('V-V','R-V','R-R') & sameIsolate == 1),
       aes(x=ds_mito,y=ds_symb,color=Groups))+
  geom_point()+
  geom_abline(slope=1, linetype=2 )+
  scale_x_continuous(limits=c(0,3),name = "Mitochondria divergence")+
  scale_y_continuous(limits=c(0,3), name="Symbionts divergence")+
  scale_color_manual(name="Holobiont pair:",labels=c("Clade I vs Clade I", "Clade I vs Clade II", "Clade II vs Clade II"),values=c('#606E96','#A5764A','#68C47F'))+
  theme_bw()

ds_plot


#### Plot as in manuscript ####

ds_plot=ggplot(data=subset(wide, Groups %in% c('V-V','R-V','R-R') &sameIsolate==1),aes(x=ds_mito,y=ds_symb))+
  # geom_point(data=subset(wide, Groups %in% c('V-V','R-V','R-R') & sameIsolate==0),size=2,shape=19,aes(color=Groups))+
  # geom_point(data=subset(wide, Groups %in% c('V-V','R-V','R-R') & sameIsolate==1),size=2,color='black',shape=21,aes(fill=Groups,shape=factor(sameIsolate)))+
    geom_point(data=subset(wide, Groups %in% c('V-V','R-V','R-R') & sameIsolate==0),aes(x=ds_mito,y=ds_symb,color=Groups,fill=Groups))+
  geom_point(data=subset(wide, Groups %in% c('V-V','R-V','R-R') & sameIsolate==1),aes(x=ds_mito,y=ds_symb,fill=Groups, shape='symb'))+
  geom_abline(slope=1, linetype=2 )+
  scale_x_continuous(limits=c(0,3),name = "Mitochondria divergence")+
  scale_y_continuous(limits=c(0,3), name="Symbionts divergence")+
  scale_fill_manual(name="Holobiont pair:",labels=c("Clade I vs Clade I", "Clade I vs Clade II", "Clade II vs Clade II"),values=c('#606E96','#A5764A','#68C47F'),guide=F)+
  scale_color_manual(name="Holobiont pair:",labels=c("Clade I vs Clade I", "Clade I vs Clade II", "Clade II vs Clade II"),values=c('#606E96','#A5764A','#68C47F'))+
    scale_shape_manual(values=c(21), labels=c('For both holobiont species\nin the pair, symbiont and\nmitochodrial genomes were\nisolated from a single individual'),name='')+
  theme_bw()+
  theme(text=element_text( family="Arial"))

ds_plot

ggsave(file='plot_dsNuclMito.svg',plot=ds_plot,width=5,height=2.665,units='in',dpi=600,device='svg')
```

# Other plots
```{r}
### Less dnds selection in the symbionts than mitochondria
dnds_plot=ggplot(subset(wide, Groups %in% c('V-V','R-V','R-R') & sameIsolate == 1),
       aes(x=dnds_mito,y=dnds_symb,color=Groups))+
  geom_point()+
  # geom_smooth(method='lm',se=F,formula = y ~ x)+
  geom_abline(slope=1, linetype=2 )+
  scale_x_continuous(limits=c(0,0.2),name = "dNdS of mitochondria")+
  scale_y_continuous(limits=c(0,0.2), name="dNdS of symbionts")+
  scale_color_manual(name="Holobiont pair:",labels=c("Gigas vs Gigas", "Gigas vs Ruthia", "Ruthia vs Ruthia"),values=c('#606E96','#A5764A','#68C47F'))+
  theme_bw()
dnds_plot

### Less purifyig selection in the symbionts than mitochondria
dn_plot=ggplot(subset(wide, Groups %in% c('V-V','R-V','R-R') & sameIsolate == 1),
       aes(x=dn_mito,y=dn_symb,color=Groups))+
  geom_point()+
  # geom_smooth(method='lm',se=F,formula = y ~ x)+
  geom_abline(slope=1, linetype=2 )+
  scale_x_continuous(limits=c(0,0.2),name = "Non-synonymous divergence of mitochondria")+
  scale_y_continuous(limits=c(0,0.2), name="Non-synonymous divergence of symbionts")+
  scale_color_manual(name="Holobiont pair:",labels=c("Gigas vs Gigas", "Gigas vs Ruthia", "Ruthia vs Ruthia"),values=c('#606E96','#A5764A','#68C47F'))+
  theme_bw()
dn_plot  

```

Are the dS in symbiont smaller than than that of mitochondria? ... YES
```{r}
levels(long$genome)
wilcox.test(ds~genome,data=long,alternative = 'greater')
?wilcox.test
```

Are the Gigas and Ruthia evolution rates (normalized by that that of their mitochondria) different? ... MAybe
```{r}
ggplot(subset(wide, Groups %in% c('V-V','R-V','R-R') & sameIsolate == 1))+
  geom_point(aes(x=Groups,y=ds_symb/ds_mito,color=Groups))

dat=subset(wide, Groups %in% c('V-V','R-R') & sameIsolate == 1)
dat$ratio=dat$ds_symb/dat$ds_mito
library(coin)
# install.packages('coin')
oneway_test(ratio~Groups,data=dat,distribution="exact") # permutation test

wilcox.test(ratio~Groups,data=dat,alternative = 'greater')

```

Difference mito/symb
```{r}
# Difference mito/symb
ggplot(wide)+
  geom_jitter(aes(x=Groups,y=ds_mito-ds_symb,color=Groups))

# quotien mito/symb
ggplot(wide)+
  geom_jitter(aes(x=Groups,y=ds_symb/ds_mito,color=Groups))

```
# dNdS vs dS
There is a very good correlation (power relation) between dnds and divergence in both mito and symbionts. 
Higher dnds in symbionts than mitoch --> More relaxed selection
```{r}
##
ggplot(long, aes(x=ds,y=dnds,color=genome))+
  geom_point(aes(shape=Groups))

ggplot(long, aes(x=ds,y=dnds,color=Groups))+
  geom_point(aes(shape=genome))

dnds_plot=ggplot(subset(wide, Groups %in% c('V-V','R-V','R-R')))+
  geom_point(aes(x=ds_mito,y=dnds_mito,color=Groups, shape='mito'))+
  geom_point(aes(x=ds_symb,y=dnds_symb,color=Groups, shape='symb'))+
  scale_x_continuous(limits=c(0,4),name = "Divergence between pairs")+
  scale_y_continuous(limits=c(0,0.25), name="Genome-wide \u03C9")+
  scale_color_manual(name="Holobiont pair:",labels=c("Gigas vs Gigas", "Gigas vs Ruthia", "Ruthia vs Ruthia"),values=c('#606E96','#A5764A','#68C47F'))+
  scale_shape_manual(values=c(21,19), labels=c('mitochondria','symbionts'),name='')+
  theme_bw()
dnds_plot

```
