---
title: "Amplicons_variance_partition"
output: html_document
editor_options: 
  chunk_output_type: console
---
# LOAD PACKAGES
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
setwd('./')
setwd("/Users/maeperez/Desktop/VesicSymb_Evolution/dowstream_analyses_and_data")
# setwd("../CRISPR_haplo/")
library(dplyr)
library(reshape2)
library(tidyr)
library(ggplot2)
library(plotly)
require(reshape)
library(devtools)
path_to_files='../mauve_alignments/bucky_results/'

```
# All genomes
## Get data
```{r}
prefix='a0.001'

nLCBs=793


pairs=read.table(paste0(path_to_files,prefix,'.pairs'),row.names=1)
head(pairs)



genes=read.table(paste0(path_to_files,prefix,'.input'),skip =2,nrows=nLCBs,sep="")
genes$V2<-strsplit(as.character(genes$V2),split="/") %>% sapply(., "[[", 2)
genes$V3<-gsub('.in','',genes$V2)
genes$V4<-substr(genes$V3,1,9)
colnames(genes)<-c('gene_no','file','nex','gene_id')

range_string = strsplit(genes$nex,split="_") %>% sapply(., "[[", 4) 
genes$len<- as.numeric(strsplit(range_string,split="to") %>% sapply(., "[[", 2)) - as.numeric(strsplit(range_string,split="to") %>% sapply(., "[[", 1))
head(genes)
nrow(genes)

dat=genes
head(dat)
nrow(dat)
summary(dat)

ggplot(dat)+
  geom_histogram(aes(x=len,y=..count..))+scale_x_log10()
```

```{r}
x=readLines(paste0(path_to_files,prefix,'.gene'))
closeAllConnections()

start <- grep("^Gene", x)
mark <- vector('integer', length(x))
mark[start] <- 1
# determine limits of each table
mark <- cumsum(mark)
# mark<-mark[1:302]
# split the data for reading
df <- lapply(split(x, mark), function(.data){
  .input <- read.table(textConnection(.data), skip=2, header=TRUE)
  attr(.input, 'name') <- .data[1]  # save the name
  .input})
# rename the list
names(df) <- sapply(df, attr, 'name')



maxtopo=setNames(data.frame(matrix(ncol = 5, nrow = 0)), c("index", "topology", "single", "joint",'gene_no'))
for(gene in names(df)){
  gene_no=gsub('Gene ','',gene) %>% substr(.,1,nchar(.)-1)

  subtable=as.data.frame(df[gene],col.names = '')
  subtable=subtable[which.max(subtable$joint),] # looking for max prob in joint
  subtable$gene_no=gene_no
  maxtopo=rbind(maxtopo,subtable)
}

colnames(maxtopo)[1]='topo_index'
nrow(maxtopo)
table=merge(dat,maxtopo,by='gene_no',all.x=T)
head(table)

topo_len=aggregate(len~topo_index, table,FUN='sum')
totlen=sum(topo_len$len)
totlen # total length (in bp) of core LCBs > 100bp
```
## Get basic stats
133 trees were necessary to represent all LCBs topologies.
The fist 10 most common topologies represent ~1/4 of LCBs and 1/2 of core genome length
```{r}
aggregate(len~topo_index, table,FUN='sum',subset = gene_id %in% subset(table,topo_index<11)$gene_id %>% unique(.) )$len %>% sum # len in bp of genes represented by fist 10 topologies

subset(table,topo_index<11)$gene_id %>% unique(.) # 217/793 LCBs (385426/762629 bp) are represented by the first 10 topologies.

# table$topo_index %>% unique

aggregate(len~topo_index, table,FUN='sum',subset = gene_id %in% subset(table,topo_index<1)$gene_id %>% unique(.) )$len %>% sum # len in bp of genes represented by fist 10 topologies

table
```

```{r}
library(ape)
rename_labels <- function(tree, vec) {
 tree$tip.label <- vec
 return(tree)
}
topo_leaves=c('Bathy','R.fausta','R.magnifica','R.pacifica', 'R.phaseoliformis','R.pliocardia','R.rectimargo','R.southwardae','SUP05','V.diagonalis','V.extenta','V.gigas1','V.gigas2','V.marissinica','V.okutanii','V.soyoae1','V.soyoae2')
names(topo_leaves)=as.character(c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17))

tree.pop<-ape::read.tree(text = '(((((1,9),((((2,7),4),5),(3,(6,8)))),(((10,11),14),(12,13))),15),16,17);')
tree.pop$tip.label<-topo_leaves[tree.pop$tip.label]
plot(root(tree.pop, c('SUP05','Bathy'))) 
  
tree.0 <- ape::read.tree(text = subset(table,topo_index==0)$topology %>% unique())
tree.0$tip.label<-topo_leaves[tree.0$tip.label]
plot(root(tree.0, c('SUP05','Bathy')))

tree.7 <- ape::read.tree(text = subset(table,topo_index==7)$topology %>% unique())
tree.7$tip.label<-topo_leaves[tree.7$tip.label]
plot(root(tree.7, c('SUP05','Bathy')))
```

```{r}
as.factor(table$topo_index) %>% unique %>% length

most.common.trees=summary(as.factor(table$topo_index)) %>% sort(.,decreasing = T) %>% data.frame(.) # sort topologies by number of represented LCBs represented by them (decreasing order)
colnames(most.common.trees)[1]='LCB_count' # # of LCB with that topology
most.common.trees$topo_index=row.names(most.common.trees)
most.common.trees$pgen=most.common.trees$LCB_count/nLCBs #pgen= LCB relative abundance
most.common.trees=most.common.trees[c(rownames(most.common.trees)[-1],rownames(most.common.trees)[1]),] # Place the "Others" category at the end as it lumps all the rare topologies


most.common.trees$cumsum=cumsum(most.common.trees$pgen) # LCB cumulative abundance

most.common.trees$n_gen=as.numeric(seq(1,nrow(most.common.trees))) # continuous axis
most.common.trees=merge(topo_len,most.common.trees,by='topo_index',all.y=T)
most.common.trees[1,'len']=totlen-sum(most.common.trees[-1,'len'])

most.common.trees=most.common.trees[c(rownames(most.common.trees)[-1],rownames(most.common.trees)[1]),] # sort dataframe again

most.common.trees$plen=most.common.trees$len/totlen
most.common.trees= most.common.trees[order(most.common.trees$plen, decreasing = T),]

rowlist=c(rownames(subset(most.common.trees, topo_index != '(Other)')),
rownames(subset(most.common.trees, topo_index == '(Other)')))
most.common.trees=most.common.trees[rowlist,]
most.common.trees$plencumsum=cumsum(as.numeric(most.common.trees$plen))


most.common.trees=most.common.trees[order(most.common.trees$n_gen),]
N50=most.common.trees[most.common.trees$cumsum>0.5,][1,'n_gen']
N75=most.common.trees[most.common.trees$cumsum>0.75,][1,'n_gen']

qplot(x=n_gen,y=cumsum,data=most.common.trees)+
  geom_vline(xintercept = N50, color='red',size=1)+
  geom_vline(xintercept = N75, color='green',size=1)+
  theme_bw() # 45 topologies are necessary to represent 75% of the LCBs, 18 topologies are necessary to represent 50% of the LCBs


most.common.trees= most.common.trees[order(most.common.trees$plencumsum),] 
most.common.trees$n_len=seq(1,nrow(most.common.trees),)

most.common.trees=most.common.trees[order(most.common.trees$n_len),]
L50=most.common.trees[most.common.trees$plencumsum>0.5,][1,'n_len']
L75=most.common.trees[most.common.trees$plencumsum>0.75,][1,'n_len']

qplot(x=n_len,y=plencumsum,data=most.common.trees) # 27 topologies are necessary to represent 75% of the core genome length, 11 topologies are necessary to represent 50% of the core genome length

most.common.trees %>% head(LCB_count,n = 20)
qplot(x=plen,y=pgen,data=most.common.trees)+scale_y_log10()+scale_x_log10()

###                                                     
ggplot(subset(table, topo_index <5))+
  geom_histogram(aes(x=topo_index+1))

# table[,c('topology','topo_index')] %>% unique()

```


A very unique sets of genes favor topology 7 and 0; For these genes, the signal for these topologies is very strong
##NMDS
```{r NMDS}
mat=as.matrix(pairs)
# set column names to gene IDs. Actual gene names would be best...
colnames(mat) = rownames(mat)
mat # Just to check. Big! 100x100 matrix
diag(mat) # check that diagonal has ones only.
# now NMDS calculations:
mds = cmdscale(1-mat)
str(mds) # matrix of 100 rows and 2 columns: for the 2 axes.
plot(mds[,1], mds[,2], type="n", xlab="axis 1", ylab="axis 2",
asp=1, # to keep aspect ratio = 1 between both axes
main="dissimilarity between genes: P{Ti and Tj differ}")
text(mds[,1], mds[,2], rownames(mds), cex=0.6)
head(mds)

mds=cbind(mds,table)
colnames(mds)[1]='NMDS1'
colnames(mds)[2]='NMDS2'
head(mds)

### Good plot ####
dev.off()

ggplot()+
    geom_point(data=subset(mds,topo_index>19),aes(x=NMDS1,y=NMDS2),color='black',size=1,alpha=0.5)+
  geom_point(data=subset(mds,topo_index<20),aes(x=NMDS1,y=NMDS2,color=as.factor(topo_index)),size=1,alpha=1)

subset(mds,topo_index==7)$topology %>% unique # Rmag at base of Clade I and V.gigas monophyletic with  Diagolanis,extenta,mariscinica
subset(mds,topo_index==0)$topology %>% unique #Rmag sister to plio,recti and V.gigas basal of Clade I
subset(mds,topo_index==1)$topology %>% unique #same as topology 0 but V.gigas monophyletic with  Diagolanis,extenta,mariscinica
subset(mds,topo_index==9)$topology %>% unique # Rmac at base of Clade II and Diagolanis,extenta at base of Clade I

#Distribution of genes with tree 0 and 1 -> genes with thes eopology tend to cluster together?
subset(mds,NMDS1>0.3)$topo_index
plot(x=subset(mds,NMDS1>0.3)$gene_id %>% strsplit(.,split = '_') %>% sapply(., "[[", 2)  %>% as.integer(.) ,y=rep(1,length(subset(mds,NMDS1>0.3)$gene_id)),type = 'h')

###
subset(mds,topo_index==0)$topo_index
plot(x=subset(mds,topo_index==0)$gene_id %>% strsplit(.,split = '_') %>% sapply(., "[[", 2)  %>% as.integer(.) ,y=rep(1,length(subset(mds,topo_index==0)$gene_id)),type = 'h')

subset(mds,topo_index==7)$topo_index
plot(x=subset(mds,topo_index==7)$gene_id %>% strsplit(.,split = '_') %>% sapply(., "[[", 2)  %>% as.integer(.) ,y=rep(1,length(subset(mds,topo_index==7)$gene_id)),type = 'h')

subset(mds,topo_index==1)$topo_index
plot(x=subset(mds,topo_index==1)$gene_id %>% strsplit(.,split = '_') %>% sapply(., "[[", 2)  %>% as.integer(.) ,y=rep(1,length(subset(mds,topo_index==1)$gene_id)),type = 'h')

subset(mds,topo_index==9)$topo_index
plot(x=subset(mds,topo_index==9)$gene_id %>% strsplit(.,split = '_') %>% sapply(., "[[", 2)  %>% as.integer(.) ,y=rep(1,length(subset(mds,topo_index==9)$gene_id)),type = 'h')

# is this random?


# distribution of all gene trees
par(mfrow = c(5, 2))
for(topo in subset(mds,topo_index<20)$topo_index %>% unique(.) %>% sort(.)){
  plot(x=subset(mds,topo_index==topo)$gene_id %>% strsplit(.,split = '_') %>% sapply(., "[[", 2)  %>% as.integer(.) ,y=rep(1,length(subset(mds,topo_index==topo)$gene_id)),type = 'h',xlab='',ylab='',sub=paste('topology', topo),ylim=c(0,1))
  
}
```

##PcoA

```{r}
library(ade4)

mat=as.matrix(pairs)
# set column names to gene IDs. Actual gene names would be best...
colnames(mat) = rownames(mat)
mat # Just to check. Big! 100x100 matrix
diag(mat) # check that diagonal has ones only.

test=as.dist(1-mat)
is.euclid(test)
is.euclid(sqrt(test))

# PcoA
cmdscale(sqrt(1-mat)) %>% plot(.)
wcmdscale(sqrt(1-mat)) %>% plot(.)

prcomp(sqrt(1-mat)) %>% plot(.)
prcomp(sqrt(1-mat)) %>% summary(.)

library(ape)
PCOA=pcoa(sqrt(test),correction = "cailliez")

barplot(PCOA$values$Relative_eig)

biplot(pcoa)


pcoa1=cmdscale(sqrt(1-mat),k=3,eig=T )
pcoa1$GOF
pcoa1$eig[1]/sum(pcoa1$eig)
pcoa1$eig[2]/sum(pcoa1$eig)
pcoa1$eig[3]/sum(pcoa1$eig)
(pcoa1$eig[1]+pcoa1$eig[2])/sum(pcoa1$eig) # axes 1 and 2 explain 10% of variation
(pcoa1$eig[1]+pcoa1$eig[2]+pcoa1$eig[3])/sum(pcoa1$eig)


pcoa1$points %>% plot(.)
pcoa=cbind(pcoa1$points,table)
colnames(pcoa)[1]='PcoA1'
colnames(pcoa)[2]='PcoA2'
colnames(pcoa)[3]='PcoA3'
head(pcoa)

### Good plot ####
ggplot()+
    geom_point(data=subset(pcoa,!(topo_index %in% c('0','1','4','6','9'))),aes(x=PcoA1,y=PcoA2,color='all others'),color='black',size=1,alpha=0.5)+
  geom_point(data=subset(pcoa,topo_index %in% c('0','1','4','6','9')),aes(x=PcoA1,y=PcoA2,color=as.factor(topo_index)),size=1,alpha=1)


plot_ly(data=subset(pcoa,topo_index %in% c('0','1','4','6','9','12','22')),x=~PcoA1,y=~PcoA2,z=~PcoA3,color=~as.factor(topo_index),size=1)

plot_ly(data=subset(pcoa,topo_index %in% c('0','1','4','6','9','12','22') & gene_no<200),x=~PcoA1,y=~PcoA2,z=~PcoA3,color=~gene_no,size=1,mode = 'lines+markers',line=(list(width=1)))
```


##MLtrees
```{r}
par(mfrow=c(1,1))

ggplot(mds)+geom_histogram(aes(x=topo_index+0.01),binwidth = 10)+scale_y_log10()

summary(as.factor(mds$topo_index)) %>% sort(.,decreasing = T) %>% plot(.,ylab='genes with that topology',xlab='topologies')

MLtrees=summary(as.factor(mds$topo_index)) %>% sort(.,decreasing = T) %>% data.frame(.) 
MLtrees=merge(unique(mds[,c('topo_index','topology')]),aggregate(mds[,c('topo_index','single','joint')],by=list(mds$topo_index),FUN = mean),by='topo_index',all.y=T) %>% .[,-3] %>% cbind(.,summary(as.factor(mds$topo_index)) %>% data.frame(.))

colnames(MLtrees)[3]='mean_single'
colnames(MLtrees)[4]='mean_joint'
colnames(MLtrees)[5]='gene_counts'
write.table(MLtrees,'MLtrees_genes20.txt',sep='\t',quote = F,row.names = F)


MLtrees$gene_counts %>% sum(.)/2


##########
most.freq.trees=summary(as.factor(mds$topo_index)) %>% sort(.,decreasing = T) %>% data.frame(.) %>% rownames(.)
most.freq.trees=most.freq.trees[1:8]

ggplot(subset(mds,topo_index %in% most.freq.trees))+
  geom_vline(aes(xintercept=gene_id %>% strsplit(.,split = '_') %>% sapply(., "[[", 2)  %>% as.integer(.),color=as.factor(topo_index)))+scale_color_brewer(palette='Dark2')


ggplot(subset(mds,topo_index %in% most.freq.trees))+
  geom_vline(aes(xintercept=gene_id %>% strsplit(.,split = '_') %>% sapply(., "[[", 2)  %>% as.integer(.),color=SEED_level_1))+facet_wrap(~topo_index)
# do these represent a random pick of the genes in the pan-genome?


### Good plot ####
most.common.trees10=most.common.trees[1:10,'topo_index']

ggplot()+
    geom_point(data=subset(mds,!(topo_index %in% most.common.trees10)),aes(x=NMDS1,y=NMDS2),color='black',size=1,alpha=0.5)+
  geom_point(data=subset(mds,topo_index %in% most.common.trees10),aes(x=NMDS1,y=NMDS2,color=as.factor(topo_index)),size=1,alpha=1)



```


# Clade II
## Get data
```{r}
prefix='a0.001_ruthia'

nLCBs=793

genes=read.table(paste0(path_to_files,prefix,'.input'),skip =2,nrows=nLCBs,sep="")
genes$V2<-strsplit(as.character(genes$V2),split="/") %>% sapply(., "[[", 2)
genes$V3<-gsub('.in','',genes$V2)
genes$V4<-substr(genes$V3,1,9)
colnames(genes)<-c('gene_no','file','nex','gene_id')

range_string = strsplit(genes$nex,split="_") %>% sapply(., "[[", 4) 
genes$len<- as.numeric(strsplit(range_string,split="to") %>% sapply(., "[[", 2)) - as.numeric(strsplit(range_string,split="to") %>% sapply(., "[[", 1))
head(genes)
nrow(genes)

dat=genes
head(dat)
nrow(dat)
summary(dat)

ggplot(dat)+
  geom_histogram(aes(x=len,y=..count..))+scale_x_log10()
```

```{r}
x=readLines(paste0(path_to_files,prefix,'.gene'))
closeAllConnections()

start <- grep("^Gene", x)
mark <- vector('integer', length(x))
mark[start] <- 1
# determine limits of each table
mark <- cumsum(mark)
# mark<-mark[1:302]
# split the data for reading
df <- lapply(split(x, mark), function(.data){
  .input <- read.table(textConnection(.data), skip=2, header=TRUE)
  attr(.input, 'name') <- .data[1]  # save the name
  .input})
# rename the list
names(df) <- sapply(df, attr, 'name')



maxtopo=setNames(data.frame(matrix(ncol = 5, nrow = 0)), c("index", "topology", "single", "joint",'gene_no'))
for(gene in names(df)){
  gene_no=gsub('Gene ','',gene) %>% substr(.,1,nchar(.)-1)

  subtable=as.data.frame(df[gene],col.names = '')
  subtable=subtable[which.max(subtable$joint),] # looking for max prob in joint
  subtable$gene_no=gene_no
  maxtopo=rbind(maxtopo,subtable)
}

colnames(maxtopo)[1]='topo_index'
nrow(maxtopo)
table=merge(dat,maxtopo,by='gene_no',all.x=T)
head(table)

topo_len=aggregate(len~topo_index, table,FUN='sum')
totlen=sum(topo_len$len)
totlen # total length (in bp) of core LCBs > 100bp
```

## Get basic stats
A total of 36 different topologies are necessary to represent all LCBs 
10 topologies are necessary to represent 75% of the LCBs, 4 topologies are necessary to represent 50% of the LCBs
```{r}
most.common.trees=summary(as.factor(table$topo_index)) %>% sort(.,decreasing = T) %>% data.frame(.) # sort topologies by number of represented LCBs represented by them (decreasing order)

colnames(most.common.trees)[1]='LCB_count' # # of LCB with that topology
most.common.trees$topo_index=row.names(most.common.trees)
most.common.trees$pgen=most.common.trees$LCB_count/nLCBs #pgen= LCB relative abundance

most.common.trees$cumsum=cumsum(most.common.trees$pgen) # LCB cumulative abundance

most.common.trees$n_gen=as.numeric(seq(1,nrow(most.common.trees))) # continuous axis
most.common.trees=merge(topo_len,most.common.trees,by='topo_index',all.y=T)

most.common.trees$plen=most.common.trees$len/totlen
most.common.trees= most.common.trees[order(most.common.trees$plen, decreasing = T),]

rowlist=c(rownames(subset(most.common.trees, topo_index != '(Other)')),
rownames(subset(most.common.trees, topo_index == '(Other)')))
most.common.trees=most.common.trees[rowlist,]
most.common.trees$plencumsum=cumsum(as.numeric(most.common.trees$plen))


most.common.trees=most.common.trees[order(most.common.trees$n_gen),]
N50=most.common.trees[most.common.trees$cumsum>0.5,][1,'n_gen']
N75=most.common.trees[most.common.trees$cumsum>0.75,][1,'n_gen']

qplot(x=n_gen,y=cumsum,data=most.common.trees)+
  geom_vline(xintercept = N50, color='red',size=1)+
  geom_vline(xintercept = N75, color='green',size=1)+
  theme_bw() # 10 topologies are necessary to represent 75% of the LCBs, 4 topologies are necessary to represent 50% of the LCBs


most.common.trees= most.common.trees[order(most.common.trees$plencumsum),] 
most.common.trees$n_len=seq(1,nrow(most.common.trees),)

most.common.trees=most.common.trees[order(most.common.trees$n_len),]
L50=most.common.trees[most.common.trees$plencumsum>0.5,][1,'n_len']
L75=most.common.trees[most.common.trees$plencumsum>0.75,][1,'n_len']

qplot(x=n_len,y=plencumsum,data=most.common.trees)+
  geom_vline(xintercept = L50, color='red',size=1)+
  geom_vline(xintercept = L75, color='green',size=1)+
  theme_bw() # 8 topologies are necessary to represent 75% of the core genome length, 3 topologies are necessary to represent 50% of the core genome length

most.common.trees %>% head(LCB_count,n = 20)
qplot(x=plen,y=pgen,data=most.common.trees)+scale_y_log10()+scale_x_log10()

###                                                     
ggplot(subset(table, topo_index <5))+
  geom_histogram(aes(x=topo_index+1))

# table[,c('topology','topo_index')] %>% unique()

```



# Clade I
## Get data
```{r}
prefix='a0.001_gigas'

nLCBs=793

genes=read.table(paste0(path_to_files,prefix,'.input'),skip =2,nrows=nLCBs,sep="")
genes$V2<-strsplit(as.character(genes$V2),split="/") %>% sapply(., "[[", 2)
genes$V3<-gsub('.in','',genes$V2)
genes$V4<-substr(genes$V3,1,9)
colnames(genes)<-c('gene_no','file','nex','gene_id')

range_string = strsplit(genes$nex,split="_") %>% sapply(., "[[", 4) 
genes$len<- as.numeric(strsplit(range_string,split="to") %>% sapply(., "[[", 2)) - as.numeric(strsplit(range_string,split="to") %>% sapply(., "[[", 1))
head(genes)
nrow(genes)

dat=genes
head(dat)
nrow(dat)
summary(dat)

ggplot(dat)+
  geom_histogram(aes(x=len,y=..count..))+scale_x_log10()
```

```{r}
x=readLines(paste0(path_to_files,prefix,'.gene'))
closeAllConnections()

start <- grep("^Gene", x)
mark <- vector('integer', length(x))
mark[start] <- 1
# determine limits of each table
mark <- cumsum(mark)
# mark<-mark[1:302]
# split the data for reading
df <- lapply(split(x, mark), function(.data){
  .input <- read.table(textConnection(.data), skip=2, header=TRUE)
  attr(.input, 'name') <- .data[1]  # save the name
  .input})
# rename the list
names(df) <- sapply(df, attr, 'name')



maxtopo=setNames(data.frame(matrix(ncol = 5, nrow = 0)), c("index", "topology", "single", "joint",'gene_no'))
for(gene in names(df)){
  gene_no=gsub('Gene ','',gene) %>% substr(.,1,nchar(.)-1)

  subtable=as.data.frame(df[gene],col.names = '')
  subtable=subtable[which.max(subtable$joint),] # looking for max prob in joint
  subtable$gene_no=gene_no
  maxtopo=rbind(maxtopo,subtable)
}

colnames(maxtopo)[1]='topo_index'
nrow(maxtopo)
table=merge(dat,maxtopo,by='gene_no',all.x=T)
head(table)
table$topo_index %>% unique # 11 different topologies in total

topo_len=aggregate(len~topo_index, table,FUN='sum')
totlen=sum(topo_len$len)
totlen # total length (in bp) of core LCBs > 100bp
```

## Get basic stats
A total of 11 different topologies are necessary to represent all LCBs 
4 topologies are necessary to represent 75% of the LCBs, 2 topologies are necessary to represent 50% of the LCBs
```{r}
most.common.trees=summary(as.factor(table$topo_index)) %>% sort(.,decreasing = T) %>% data.frame(.) # sort topologies by number of represented LCBs represented by them (decreasing order)
colnames(most.common.trees)[1]='LCB_count' # # of LCB with that topology
most.common.trees$topo_index=row.names(most.common.trees)
most.common.trees$pgen=most.common.trees$LCB_count/nLCBs #pgen= LCB relative abundance

# most.common.trees=most.common.trees[c(rownames(most.common.trees)[-1],rownames(most.common.trees)[1]),] # Place the "Others" category at the end as it lumps all the rare topologies


most.common.trees$cumsum=cumsum(most.common.trees$pgen) # LCB cumulative abundance

most.common.trees$n_gen=as.numeric(seq(1,nrow(most.common.trees))) # continuous axis
most.common.trees=merge(topo_len,most.common.trees,by='topo_index',all.y=T)

most.common.trees$plen=most.common.trees$len/totlen
most.common.trees= most.common.trees[order(most.common.trees$plen, decreasing = T),]

rowlist=c(rownames(subset(most.common.trees, topo_index != '(Other)')),
rownames(subset(most.common.trees, topo_index == '(Other)')))
most.common.trees=most.common.trees[rowlist,]
most.common.trees$plencumsum=cumsum(as.numeric(most.common.trees$plen))


most.common.trees=most.common.trees[order(most.common.trees$n_gen),]
N50=most.common.trees[most.common.trees$cumsum>0.5,][1,'n_gen']
N75=most.common.trees[most.common.trees$cumsum>0.75,][1,'n_gen']

qplot(x=n_gen,y=cumsum,data=most.common.trees)+
  geom_vline(xintercept = N50, color='red',size=1)+
  geom_vline(xintercept = N75, color='green',size=1)+
  theme_bw() # 4 topologies are necessary to represent all of the LCBs, 2 topologies are necessary to represent 50% of the LCBs


most.common.trees= most.common.trees[order(most.common.trees$plencumsum),] 
most.common.trees$n_len=seq(1,nrow(most.common.trees),)

most.common.trees=most.common.trees[order(most.common.trees$n_len),]
L50=most.common.trees[most.common.trees$plencumsum>0.5,][1,'n_len']
L75=most.common.trees[most.common.trees$plencumsum>0.75,][1,'n_len']

qplot(x=n_len,y=plencumsum,data=most.common.trees)+
  geom_vline(xintercept = L50, color='red',size=1)+
  geom_vline(xintercept = L75, color='green',size=1)+
  theme_bw() # 4 topologies are necessary to represent 75% of the core genome length, 2 topologies are necessary to represent 50% of the core genome length

most.common.trees %>% head(LCB_count,n = 20)
qplot(x=plen,y=pgen,data=most.common.trees)+scale_y_log10()+scale_x_log10()

###                                                     
ggplot(subset(table, topo_index <5))+
  geom_histogram(aes(x=topo_index+1))

# table[,c('topology','topo_index')] %>% unique()

```


